Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2017-07-05 paper.tex
edits to domestication section

\section*{Re-evaluating concepts of domestication}

@@

\subsection*{What is the progenitor of a crop?}
Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or completely confounded.
The level of divergence of a crop from newly encountered populations and species will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
Similarly, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
Recent work has shown that, following the original domestication event in the central Andes from \emph{Solanum tuberosum}, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013}.
Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a reevaluation of crop origins.
In cases like maize and potato it is important to recognize the substantial contributions of introgressing taxa to the genetic base of modern crops.
Broad recognition of the role these wild relatives have played in crop adaptation could further their use in breeding and elevate their conservation status.

\subsection*{When was a crop domesticated?}
Estimates of the timing of domestication based on levels of sequence divergence may be affected when introgressed haplotypes are included.
The directionality of this effect is likewise dependent on $N_e$ of the donor population.

\subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
Estimates of the initial domestication bottleneck may be skewed when introgression is not considered.
Chromosomal regions experiencing introgression may have an altered effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
For example, introgression from wild taxa with historically high $N_e$ will lead to underestimates of the strength of the domestication bottleneck.
Conversely, if the donor population has a relatively small effective population size, the opposite bias may be imposed upon bottleneck estimations.
In most cases, the effective population size of the domesticated crop will be lower than that of the wild progenitor and wild relative populations.

>>>>

\section*{Re-evaluating domestication}

@@

\subsection*{What is the progenitor of a crop?}
Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or completely confounded.
The level of divergence of a crop from newly encountered populations and species will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining that widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
Similarly, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013}.
Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a reevaluation of crop origins.
In cases like maize and potato it is important to recognize the substantial contributions of introgressing taxa to the genetic base of modern crops.
Broad recognition of the role these wild relatives have played in crop adaptation could further their use in breeding and elevate their conservation status.

\subsection*{When was a crop domesticated?}
Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
In highly introgressed domesticates, these estimates will be based on comparison of either crop or introgressant haplotypes to those of the progenitor.
In such cases, divergence is a mixture of time since domestication and time since split of the progenitor and introgressing species.
This phenomenon, in combination with divergence of samples from true ancestral populations and ongoing evolution of crop progenitors, may explain discrepancies between domestication dates based on genetic and archaeological data. \mbh{perhaps add citations here}
More accurate genetic estimates of the timing of domestication may be obtained by excluding loci that show signatures of introgression.

\subsection*{How was genome-wide diversity impacted by a domestication bottleneck?}
Estimates of the strength of the initial domestication bottleneck may also be skewed when introgression is not considered.
Introgressed chromosomal regions may have a distinct effective population size ($N_e$) relative to non-introgressed regions depending on diversity within the donor taxon.
The influence of introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of introgression, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
For example, substantial introgression at neutral loci from wild taxa with a historically higher $N_e$ will lead to underestimates of the overall strength of the domestication bottleneck.
